---
    title: "TP2"
    output: html_document
---

``` {r include = FALSE}
# Load libraries
library("expm")
```

# Basic PageRank

## Functions

These are the functions we created to compute the PageRank index.

### Sum powers of a matrix

Computes the sum of n first powers of matrix m.

n.d is a damping factor applied to the powers of m.

Example : *sum.powers.matrix(m, 3, 2)* returns: $m + \frac{1}{2^2} * m^2 + \frac{1}{3^2} * m^3$

``` {r}
sum.powers.matrix <- function(m, n, n.d) {
    powers <- c(1:n)
    res <- Reduce('+', lapply(powers, function(x) (1 / (x ^ n.d)) * (m %^% x)))

    return(res)
}
```
### Remove auto-references

Removes all the auto-references in matrix m (ie puts the diagonal to 0).

``` {r}
remove.autoreferences <- function(m) {
    res <- m
    diag(res) <- 0

    return(res)
}
```
### PageRank iteration

Computes an iteration of the PageRank algorithm.

Parameters:

  - refs: the references matrix
  - n: the number of powers of refs to consider (ie the depth of references)
  - d: the PageRank damping factor
  - n.d: the damping factor for powers of refs (see function sum.powers.matrix)
  - pr: the current PageRank values

``` {r}
pagerank.iteration <- function(refs, n, d, n.d, pr) {
    # Number of articles
    n.articles <- dim(refs)[1]

    # n-level references
    m <- sum.powers.matrix(refs, n, n.d)

    # remove auto-references
    m <- remove.autoreferences(m)

    # Compute PageRank
    pr.res <- (1-d)/n.articles + (d * (m %*% (pr/colSums(m))))

    return(pr.res)
}

```

## PageRank flow

Read source file and turn it into a matrix (for later work).

``` {r}

# Read source file
data <- read.table("citeseer.rtable")

# Cast data to matrix
references <- as.matrix(data)

```

Example of PageRank computations with a simple reference matrix.

``` {r}
m <- matrix(c(0,1,1,0,0,1,1,0,0),3)
```

``` {r echo = FALSE}
print(m)
```

Initial PageRank values:

``` {r}
pr <- rep(1,3)
```

``` {r echo = FALSE}
print(pr)
```

3 PageRank iterations:

``` {r}

pr <- pagerank.iteration(m, 2, 0.85, 2, pr)
pr <- pagerank.iteration(m, 2, 0.85, 2, pr)
pr <- pagerank.iteration(m, 2, 0.85, 2, pr)

```

``` {r echo = FALSE}
print(pr)
```
