---
title: "TP2"
output:
  pdf_document: default
  html_document: default
---

``` {r include = FALSE}
# Load libraries
library("expm")
```

#PageRank
##
Olivier Sirois, 1626107 et 
Alexandre Bento, _______

## Functions

These are the functions we created to compute the PageRank index.

### Sum powers of a matrix

Computes the sum of n first powers of matrix m.

n.d is a damping factor applied to the powers of m.

Example : *sum.powers.matrix(m, 3, 2)* returns: $m + \frac{1}{2^2} * m^2 + \frac{1}{3^2} * m^3$

``` {r}
sum.powers.matrix <- function(m, n) {
  powers <- c(1:n)
  res <- Reduce('+', lapply(powers, function(x) m %^% x))
  res[res > 1] <- 1
  
  return(res)
}
```
### Remove auto-references

Removes all the auto-references in matrix m (ie puts the diagonal to 0).

``` {r}
remove.autoreferences <- function(m) {
    res <- m
    diag(res) <- 0

    return(res)
}
```


### PageRank iteration

Computes an iteration of the PageRank algorithm.

Parameters:

  - refs: the references matrix
  - n: the number of powers of refs to consider (ie the depth of references)
  - d: the PageRank damping factor
  - n.d: the damping factor for powers of refs (see function sum.powers.matrix)
  - pr: the current PageRank values



```{r}
#Cette fonction fait une itérations de l'algorithme page rank. Nous avons ajouter une petite modifications dans le cas ou la somme d'un colonne est de 0 (c'est à dire, une 
# article non référencé) on remplace ensuite la valeur qui sera égale a Inf, par un 0
page.rank <- function(m, d, pr){
  denum <- (pr/colSums(m))
  denum[denum == Inf] <- 0
  (pr <- (1-d)/3 + (d * (m %*%denum)))
  #print(pr[450])
  return(pr)
}
```
###corr.vm

We're also using some of the other functions used in TP1 in regards to item-item recommandations.

corr.vm Computes the Pearson correlations between the different rows of a matrix.

```{r}
#Correlation entre la rangee v (v = index) et chaque colonne de la matrice m
corr.vm <- function(v,m) {
  # on centre les valeurs de la matrice m en fonction de la moyenne, on la renomme m.centre
  v.i <- rowMeans(m[,], na.rm=T)
  # on enleve les NA
  m[is.na(m)] <- 0
  m.centre <- m - v.i
  # on centre le vecteur v en fonction de sa moyenne
  v.index <- v
  v.index[is.na(v)] <- 0
  v.index <- v.index - mean(v, na.rm=T)
  # on retourne ensuite un vecteur correspondant entre le vecteur v et sa correlation avec chaque rangers de m
  return( (v.index%*%t(m.centre))/(sqrt(sum(v.index^2) * rowSums(m.centre^2))))
}
```
###cosinus.vm
function that computes the Cos between all the different columns of a matrix
```{r}
## Cosinus entre un vecteur v et chaque colonne dela matrice m
cosinus.vm <- function(v,m) {
  # On on met tous nos valeurs de NA a 0, sinon on va avoir des problemes de calculs avec des matrices sparse
  m[is.na(m)] <- 0
  v[is.na(v)] <- 0 ;
  # On calcule le cosinus entre le vecteur V et les colonnes de la matrice m en utilisant la formule vu en classe
  (v %*% m)/(sqrt(colSums(m^2)) * sqrt(sum(v^2))) 
}
```
###page.rank.until.stab
This function calls an iteration of page rank until our rankings are stabilized. Stabilized meaning that the mean error between our last two iterations of page rank is less then 0.0001

```{r}
# Cette fonction exécute l'algorithme PageRank jusqu'à ce que c'est valeurs soit stabilisés. On définie une stabilit. lorsque l'erreur moyenne absolue est moins de .0001 
page.rank.until.stab <- function(m, d, pr){
  pr.next <- page.rank(m, d, pr)
  while(abs(mean(pr.next-pr)) > 0.0001){
    pr <- pr.next
    pr.next <- page.rank(m,d,pr)
  }
  return (pr)
}
```

## PageRank flow

Read source file and turn it into a matrix (for later work).

```{r}

# Read source file
data <- read.table("citeseer.rtable")

# Cast data to matrix
references <- as.matrix(data)

```

Example of PageRank computations with a simple reference matrix.

```{r}
m <- matrix(c(0,1,1,0,0,1,1,0,0),3)
```

``` {r echo = FALSE}
print(m)
```

Initial PageRank values:

```{r}
pr <- rep(1,3)
```

```{r echo = FALSE}
print(pr)
```

3 PageRank iterations:

```{r}
d <- 0.85
pr <- page.rank(m, d,  pr)
print(pr)
pr <- page.rank(m, d,  pr)
print(pr)
pr <- page.rank(m, d,  pr)

```

``` {r echo = FALSE}
print(pr)
```


#Question 1
We start our first question by importing all our data from our citeseer dataset.

```{r}
m = as.matrix(read.table("citeseer.rtable", check.names=F))
library("expm")
library("dplyr")
d <- 0.85
pr <- rep(1, dim(m)[1])
```

We then calculate the domain of our search. We specify $S_{primary}$ as our primary domain being the references of our target article, which is 422908.
We're calculating our extended domain, which is $S'$ and defined as the domain of all our articles in $S$. Which can be described as $S' = S \wedge \sum_i S(S_i)$. We calculate $S'$ by squaring our references matrix, which gives us the references of our references and then adding it to our previous matrix. All values for an article $x$ that are higher then 0 are considered in our domain.

```{r}
# On calcule notre domaine S comme étant tout l
#es article sont référencés par notre origine. Pour faire cela,
#on regarde ceux qui on une valeur positive dans notre matrice référentielle
S <- which(m["422908",]==1)
# Pour le domaine S prime, nous voulons aussi rajouter les références des références. 
#Pour faire cela, nous faisont que prendre la somme des deux première puissance de la
# matrice référentielle. C'est a dire, la matrice référentielle elle-même 
#et la deuxième puissance.
m.prime <- sum.powers.matrix(m,2)

#on enleve les auto-references
diag(m.prime) <- 0
#et on prend les valeurs positives
S.prime <- which(m.prime["422908",]==1)

print(S)
print(S.prime)
```

We then continue with on with calculating the PageRank of all our articles. We do this globally and locally.

```{r}
#On calcule le pagerank de tout nos articles (pas très long)
pr <- page.rank.until.stab(m, d, pr)

pr.S <- rep(1, dim(m[S,S])[1])
#on calcule le page rank local du domaine S
pr.S <- page.rank.until.stab(m[S,S], d, pr.S)
#et apres pour S prime
pr.S.prime <- rep(1, dim(m[S.prime,S.prime])[1])

pr.S.prime <- page.rank.until.stab(m[S.prime, S.prime], d, pr.S.prime)
```

And then we extract from all the Page Ranks, the one that we are interested in, being the ones in the $S_{primary}$ domain and the $S'$ domain. We only do this in the global PageRanks since all the data from the local PageRanks are to be used for recommendations.

```{r}
#on place les rankings PageRank de notre domaine dans un vecteur
S.rankings <- pr[S]
#pour le domaine S prime
S.prime.rankings <- pr[S.prime]
#On crée un dataframe avec nos données
S.dat <- data.frame(S.rankings, S)
S.dat$article = rownames(S.dat)
#même chose pour le domaine S prime 
S.prime.dat <- data.frame(S.prime.rankings, S.prime)
S.prime.dat$article = rownames(S.prime.dat)
print.data.frame(S.prime.dat)
print.data.frame(S.dat)
```

after sorting all the values, we can determine which ones we will recommend.

```{r}
#On crée un dataframe avec nos données
S.loc.dat <- data.frame(pr.S)
S.loc.dat$article <- rownames(pr.S)
#meme chose pour S prime (local)
S.prime.loc.dat <- data.frame(pr.S.prime)
S.prime.loc.dat$article <- rownames(pr.S.prime)
#on trie pour S et S prime
S.loc.best <- S.loc.dat %>% select(pr.S, article) %>% arrange(desc(pr.S, arr.ind=T))
S.prime.loc.best <- S.prime.loc.dat %>% select(pr.S.prime, article) %>% arrange(desc(pr.S.prime, arr.ind=T))

#on effectue un trie sur nos valeurs, on sort celles qui sont les plus hautes en premier
S.best <- S.dat %>% select(S.rankings, S, article) %>% arrange(desc(S.rankings, arr.ind=T))
#même chose pour le domaine S prime
S.prime.best <- S.prime.dat %>% select(S.prime.rankings, S.prime, article) %>% arrange(desc(S.prime.rankings, arr.ind=T))
print("Version Globales")
head(S.best)
head(S.prime.best)
print("Versions Locales")
head(S.loc.best)
head(S.prime.loc.best)
```

#Question 2
For the second question, we will compute recommendations based on the same principles then the TP1. We will use and item-item approch and recommend articles based on the similarity of their references. To do that, we will simply compute the Pearson Correlation and the Cosine between all of their rows.


```{r}
## on calcule nos coefficients de correlation et du cosinus
corr.ratings <- corr.vm(m["422908", ], m[S,])
cos.ratings <- cosinus.vm(m["422908",], t(m[S,]))

# on remplace les NaN par 0
corr.ratings[is.nan(corr.ratings)] <- 0
cos.ratings[is.nan(cos.ratings)] <- 0

# on prends nos etiquettes
labels.corr <- colnames(corr.ratings)
labels.cos <- colnames(cos.ratings)

# on cree nos dataframes
df.corr <- data.frame(corr = as.vector(corr.ratings), article = labels.corr)
df.cos <- data.frame(cos = as.vector(cos.ratings), article = labels.cos)
```

Having done that, we can sort all of our coefficients and then take our highest ones for recommendation.

```{r}
#on trie
df.best.cos <- df.cos %>% select(cos, article) %>% arrange(desc(cos, arr.ind=T))
df.best.corr <- df.corr %>% select(corr, article) %>% arrange(desc(corr, arr.ind=T))

print.data.frame(df.best.cos)
print.data.frame(df.best.corr)
```

#Interpretation
Judging from the recommendation of all our possible scenarios. We can safely assume that all of our recommendations are admissible. Admissible being that they are related to the original article in question. 

We can see that the article 155792 is one of the few that figures in all our recommendations. It is recommended 3rd by both our item-item approaches and is recommended by the two domains of the PageRanks (global and local) approaches. This articles is about logics and models of real time, while the original article is about symbolic model checking for real-time systems. This is a very good recommendation.

It is also a good time to mention the differences observe between the Local PageRanks and the Global PageRanks. The difference being that the computation of the scores for the local versions are done on only the domain specifically whilst the global version is done on all the articles of the matrix. Intuitively, it should yield similar results. Although in reality, it does not. We can see that the recommandations based on $S$ and $S'$ are almost exactly the same while the scores on the global version of $S$ and $S'$ are very different. This is due to the fact that the local version doesn't take into account the references of article that aren't necessarily in the domain. This means that the local version of PageRanks scores on the popularity of the article in the domain only, while the global version scores on the popularity versus all of the article in the database but we only look at the articles in the domain for recommendation.

Now, if we look at our top recommendation of our $S'$ domain of the global, we can see that the recommendation is a little bit different. The article recommended is about minimal state graph generation. It is recommending this article because while not being popular in the domain, it is a very general topic and very useful article in all sorts of disciplines thus demonstrating an advantage of the global pagerank versus the local pagerank. While not being closely related to real-time systems, it might be relevant through the mathematical principles found in the article itself and may be a very good recommendation based on what the reader is actually trying to accomplish.

On the other hand, The top recommendations of our item-item approaches are still about real-time systems while being part of our domains $S$ and by extent $S'$

#Conclusion

Judging from our results, we can conclude that both methods yield quite powerful results. Though they are very different in nature..

Our item-item approach gives recommendations that are very close in terms of euclidian distance. They account for the references between the articles and is not subject to a restricted domain. They look at the reference matrix in its whole for articles that are close to each other and calculates the ones that are closer to each other. Judging from our results, we can safely say that this method can be used for recommendations purposes.

the Page Rank approach gives a much more personnalized approach. the page rank algorithm is inherently more axed toward the relations between the articles thus making it more personalized then the item-item approach. One of the key factors of the Page Rank is the establishment of the domain. In our experiment, we used $S$ and $S'$  as the domains for our algoriths. Both of these domains yielded great results (Results that are very similar to the item-item approach), but $S'$ yielded more interesting results. The top recommendation for $S'$ isn't closely related to the original article itself, but it explains one particular part of the article in better detail. This makes it a better recommendation by nature then most recommendations given by $S$ and the item-item approaches.

Although we concluded that we had good recommendations, it is ultimately depending on the user. What we consider a good recommendation might be considered a bad recommendation by a user who is looking into writing his research paper while our recommendation is a perfect recommendation for someone who is looking for some light reading. It depends on what the user needs.








